---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# drep

<!-- badges: start -->
<!-- badges: end -->

`drep` allows calculation of the Dengue Reproduction number, $R_0$, from 
Force of Infection using the at-equilibrium number of primary, secondary, tertiary 
and quaternary infections in a population and their relative infectiousness. 
This is done assuming that Dengue transmission is at equilibrium. 

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mrc-ide/drep")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(drep)
## basic example code
```

Define some parameters

* number of age groups in the country human population (`n_age_groups`)
* total human population of the country (`pop`)
* the Force of Infection estimate (`FOI`)
* relative infectiousness of the four dengue infections (`phis`)

```{r}
n_age_groups <- 20
pop <- 500000
FOI <- 0.0235
phis <- c(1, 1, 1, 1)
```

Simulate some data

* lower and upper age limits (`l_lim` and `u_lim`) of the country age groups 
* number of individuals in each age group (`n_j`) 
* proportion of individuals in each age gorup (`f_j`)

```{r}
l_lim <- seq(0, 95, length.out = n_age_groups)
u_lim <- seq(5, 100, length.out = n_age_groups)
n_j <- sample(1:50, n_age_groups, replace = TRUE)
f_j <- n_j / sum(n_j)
```

Calculate the $R_0$

```{r}
R0 <- calculate_R0(FOI, pop, f_j, u_lim, l_lim, phis)
R0
```
