---
title: "Burden estimation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{burden}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(drep)
```

## Introduction

This vignette explains how to use `drep` to translate an estimate of the force of infection of dengue to the corresponding number of infections, symptomatic infections ($_cases_$) and symptomatic infections requiring hospitalization ($_hospitalized_cases_$). To do this we need:

The `population` of the data point will also be needed when calculating the $R_0$.

We first consider the fourth data point from the `foi` dataset (see [vignette](https://mrc-ide.github.io/drep/articles/burden.html)), i.e. a `caseReport` force of infection estimate from Brazil:

```{r}
lambda <- drep::foi[4, "FOI"]
lambda
```

We extract the population of that data point:

```{r}
id_0 <- drep::foi[4, "ID_0"]
pop_size <- drep::foi[4, "population"]
pop_size
```

We then define some paramaters:

* the lower and upper age limits (`l_lim` and `u_lim`) of the country age groups;
* the proportions of primary, secondary, and tertiary/quaternary infections which are symptomatic (`sym_weights`);
* the proportions of primary, secondary, and tertiary/quaternary cases which require hospitalization (`hosp_weights`);

```{r}
n_age_groups <- 20
l_lim <- seq(0, 95, length.out = n_age_groups)
u_lim <- seq(5, 100, length.out = n_age_groups)
sym_weights <- c(0.45, 0.85, 0.15)
hosp_weights <- c(0.04, 0.1, 0.04)
```

We can then calculate the total annual number of infections which correspnds to `lambda`:

```{r}
n_infections <- calculate_infections(lambda, n_j, u_lim, l_lim, pop_size)
n_infections
```

the number of cases:

```{r}
n_cases <- calculate_cases(lambda, n_j, u_lim, l_lim, sym_weights, pop_size)
n_cases
```

and the number of hopsitalized cases:

```{r}
n_h_cases <- calculate_hosp_cases(lambda, n_j, u_lim, l_lim, hosp_weights, sym_weights, pop_size)
n_h_cases
```
