---
title: "Estimating reproduction number and burden of Dengue"
author: "Lorenzo Cattarino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimating reproduction number and burden of Dengue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Background

Accurate estimates of transmission intensity of Dengue are needed to quantify the global burden of the disease and assess the impact of control strategies. Transmission intensity is often measured using the force of infection (the per capita rate at which susceptible individuals acquire infection, **$\lambda$**) or the basic reproduction number (the average number of secondary infections generated by one primary case entering a susceptible population, **R~0~**). Although they measure different things, **$\lambda$** and **R~0~** are related through the age structure of the population affected by the disease. This package provides a set of functions to translate force of infection into reproduction number and to quantify number of infections and symptomatic cases. We make a number of assumptions: 

1. Dengue transmission is at equilibrium 
2. There is a maxiumum number of four dengue infections per individual
3. Force of infection is constant through time
4. Four dengue serotypes in circulation 

# Getting started

## Installation

You can install the development version of the package using `devtools`. First install `devtools`, if you don't already have it

```{r include = TRUE, eval = FALSE}
install.packages("devtools")
library(devtools)
```

Then, in a fresh R session, install the `drep` package

```{r include = TRUE, eval = FALSE}
devtools::install_github("mrc-ide/drep", ref = "mrc-branch")
library(drep)
```

We will start with a simple example. The `foi` dataset contains force of infection estimates for different locations in the world. It looks like this

```{r, echo = FALSE, results = 'asis'}
knitr::kable(head(mtcars, 10))
```

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(foi, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
